---
title: "Pokemon EDA - Exploratory Data Analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: cosmo
  # gihub_document:
  #   toc: true
  #   toc_depth: 2
  #   number_sections: true
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "reports"
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

## 1. Load Libraries and Data

```{r libraries}
library(tidyverse)
library(ggplot2)
library(knitr)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))

# Check for optional packages
has_corrplot <- requireNamespace("corrplot", quietly = TRUE)
has_patchwork <- requireNamespace("patchwork", quietly = TRUE)
has_kableExtra <- requireNamespace("kableExtra", quietly = TRUE)

if (has_patchwork) library(patchwork)
```

```{r load-data}
# Load cleaned data
pokemon_df <- read_csv("data/pokemon_data.csv")

# Display structure
glimpse(pokemon_df)
```

## 2. Feature Engineering: Labels & Categories

### 2.1 Category Labels

```{r add-labels}
# Legendary Pokemon
legend_dex <- c(
  144, 145, 146, 150,
  243, 244, 245, 249, 250,
  377, 378, 379, 380, 381, 382, 383, 384,
  480, 481, 482, 483, 484, 485, 486, 487, 488,
  638, 639, 640, 641, 642, 643, 644, 645, 646,
  716, 717, 718,
  772, 773, 785, 786, 787, 788, 789, 790, 791, 792,
  800, 888, 889, 890, 891, 892, 894, 895, 896, 897, 898, 905,
  1001, 1002, 1003, 1004, 1007, 1008, 1014, 1015, 1016, 1017, 1024
)

# Mythical Pokemon
myth_dex <- c(
  151, 251, 385, 386, 489, 490, 491, 492, 493,
  494, 647, 684, 689, 719, 720, 721,
  801, 802, 807, 808, 809, 893, 1025
)

# Paradox Pokemon
paradox_dex <- c(984:995, 1005:1010, 1020:1023)

# Ultra Beast Pokemon
ub_dex <- c(793:799, 803:806)

# Add category columns
pokemon_df <- pokemon_df |>
  mutate(
    is_legendary = dex %in% legend_dex,
    is_mythical = dex %in% myth_dex,
    is_paradox = dex %in% paradox_dex,
    is_ultra_beast = dex %in% ub_dex,
    # Combined special category
    is_special = is_legendary | is_mythical | is_paradox | is_ultra_beast,
    # Category label for plotting
    category = case_when(
      is_legendary ~ "Legendary",
      is_mythical ~ "Mythical",
      is_paradox ~ "Paradox",
      is_ultra_beast ~ "Ultra Beast",
      TRUE ~ "Regular"
    )
  )
```

### 2.2 Generation Assignment

```{r add-generation}
pokemon_df <- pokemon_df |>
  mutate(
    generation = case_when(
      dex <= 151 ~ "Gen 1",
      dex <= 251 ~ "Gen 2",
      dex <= 386 ~ "Gen 3",
      dex <= 493 ~ "Gen 4",
      dex <= 649 ~ "Gen 5",
      dex <= 721 ~ "Gen 6",
      dex <= 809 ~ "Gen 7",
      dex <= 905 ~ "Gen 8",
      TRUE ~ "Gen 9"
    )
  )

# Set factor order
pokemon_df$generation <- factor(pokemon_df$generation,
  levels = paste0("Gen ", 1:9)
)
pokemon_df$category <- factor(pokemon_df$category,
  levels = c(
    "Regular", "Legendary", "Mythical",
    "Paradox", "Ultra Beast"
  )
)
```

### 2.3 Type One-Hot Encoding

```{r type-encoding}
# Get all unique types
all_types <- sort(unique(na.omit(c(pokemon_df$type_1, pokemon_df$type_2))))

cat("Unique types:", paste(all_types, collapse = ", "), "\n")

# Create one-hot encoding for each type
for (type in all_types) {
  col_name <- paste0("has_", tolower(type))
  pokemon_df[[col_name]] <- (pokemon_df$type_1 == type | pokemon_df$type_2 == type)
}

# Count dual-type Pokemon
pokemon_df <- pokemon_df |>
  mutate(is_dual_type = !is.na(type_2))
```

## 3. Data Overview

### 3.1 Basic Statistics

```{r summary-stats}
# Numeric columns
num_cols <- c(
  "total", "hp", "attack", "defense", "sp_atk", "sp_def", "speed",
  "height_m", "weight_kgs", "bmi"
)

# Summary table
summary_stats <- pokemon_df |>
  select(all_of(num_cols)) |>
  summary()

print(summary_stats)
```

### 3.2 Category Counts

```{r category-counts}
# Count by category
category_summary <- pokemon_df |>
  count(category) |>
  mutate(percentage = round(100 * n / sum(n), 2)) |>
  arrange(desc(n))

print(kable(category_summary, caption = "Pokemon by Category"))

# Count by generation
gen_summary <- pokemon_df |>
  count(generation) |>
  mutate(percentage = round(100 * n / sum(n), 2))

print(kable(gen_summary, caption = "Pokemon by Generation"))

# Type distribution (primary type)
type_summary <- pokemon_df |>
  count(type_1, sort = TRUE) |>
  mutate(percentage = round(100 * n / sum(n), 2))

print(kable(head(type_summary, 10), caption = "Pokemon by Primary Type (Top 10)"))
```

## 4. Univariate Analysis

### 4.1 Base Stat Total Distribution

```{r total-dist}
# Overall distribution
p1 <- ggplot(pokemon_df, aes(x = total)) +
  geom_histogram(bins = 40, fill = "steelblue", alpha = 0.7) +
  geom_vline(aes(xintercept = mean(total, na.rm = TRUE)),
    color = "red", linetype = "dashed", size = 1
  ) +
  labs(
    title = "Distribution of Base Stat Total (BST)",
    subtitle = paste("Mean:", round(mean(pokemon_df$total, na.rm = TRUE), 1)),
    x = "Total BST", y = "Count"
  )

# By category
p2 <- ggplot(pokemon_df, aes(x = total, fill = category)) +
  geom_density(alpha = 0.5) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "BST Distribution by Category",
    x = "Total BST", y = "Density", fill = "Category"
  )

if (has_patchwork) {
  p1 / p2
} else {
  print(p1)
  print(p2)
}
```

### 4.2 Individual Stat Distributions

```{r stat-distributions}
stat_cols <- c("hp", "attack", "defense", "sp_atk", "sp_def", "speed")

pokemon_long <- pokemon_df |>
  select(dex, name, category, all_of(stat_cols)) |>
  pivot_longer(
    cols = all_of(stat_cols),
    names_to = "stat",
    values_to = "value"
  )

ggplot(pokemon_long, aes(x = value, fill = stat)) +
  geom_histogram(bins = 30, alpha = 0.7) +
  facet_wrap(~stat, scales = "free_y") +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title = "Distribution of Individual Base Stats",
    x = "Stat Value", y = "Count"
  ) +
  theme(legend.position = "none")
```

### 4.3 Physical Attributes

```{r physical-dist}
p1 <- ggplot(pokemon_df, aes(x = height_m)) +
  geom_histogram(bins = 40, fill = "darkgreen", alpha = 0.7) +
  scale_x_log10() +
  labs(
    title = "Height Distribution (log scale)",
    x = "Height (m)", y = "Count"
  )

p2 <- ggplot(pokemon_df, aes(x = weight_kgs)) +
  geom_histogram(bins = 40, fill = "darkorange", alpha = 0.7) +
  scale_x_log10() +
  labs(
    title = "Weight Distribution (log scale)",
    x = "Weight (kg)", y = "Count"
  )

p3 <- pokemon_df |>
  filter(bmi < 1000) |> # Filter extreme outliers (Cosmoem BMI=99990)
  # otherwise the plot will be severly compressed
  ggplot(aes(x = bmi)) +
  geom_histogram(bins = 40, fill = "purple", alpha = 0.7) +
  labs(
    title = "BMI Distribution (filtered < 1000)",
    subtitle = "Note: Cosmoem (BMI=99,990) excluded as extreme outlier",
    x = "BMI", y = "Count"
  )

if (has_patchwork) {
  (p1 | p2) / p3
} else {
  print(p1)
  print(p2)
  print(p3)
}
```

## 5. Bivariate Analysis

### 5.1 Correlation Matrix

```{r correlation}
# Compute correlation
cor_matrix <- pokemon_df |>
  select(all_of(num_cols)) |>
  cor(use = "pairwise.complete.obs")

# Visualize
if (has_corrplot) {
  corrplot::corrplot(cor_matrix,
    method = "color",
    type = "upper",
    tl.col = "black",
    tl.srt = 45,
    addCoef.col = "black",
    number.cex = 0.7,
    title = "Correlation Matrix of Pokemon Attributes",
    mar = c(0, 0, 2, 0)
  )
} else {
  # Alternative: heatmap
  cor_df <- as.data.frame(cor_matrix) |>
    rownames_to_column("var1") |>
    pivot_longer(-var1, names_to = "var2", values_to = "correlation")

  ggplot(cor_df, aes(x = var1, y = var2, fill = correlation)) +
    geom_tile() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
    labs(title = "Correlation Matrix of Pokemon Attributes") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

### 5.2 Attack vs Defense

```{r attack-defense}
ggplot(pokemon_df, aes(x = attack, y = defense, color = category)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "Attack vs Defense by Category",
    x = "Attack", y = "Defense", color = "Category"
  )
```

### 5.3 Special Attack vs Special Defense

```{r sp-attack-defense}
ggplot(pokemon_df, aes(x = sp_atk, y = sp_def, color = category)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "Special Attack vs Special Defense by Category",
    x = "Special Attack", y = "Special Defense", color = "Category"
  )
```

### 5.4 Height vs Weight

```{r height-weight}
ggplot(pokemon_df, aes(x = height_m, y = weight_kgs, color = type_1)) +
  geom_point(alpha = 0.6, size = 2) +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(
    title = "Height vs Weight (log-log scale)",
    x = "Height (m)", y = "Weight (kg)", color = "Primary Type"
  ) +
  theme(legend.position = "right")
```

## 6. Group Comparisons

### 6.1 Stats by Category

```{r stats-by-category}
# Box plots for each stat
pokemon_long |>
  ggplot(aes(x = category, y = value, fill = category)) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(~stat, scales = "free_y") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Base Stats by Pokemon Category",
    x = "Category", y = "Stat Value"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

### 6.2 BST by Generation

```{r bst-generation}
ggplot(pokemon_df, aes(x = generation, y = total, fill = generation)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.2, alpha = 0.5) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    title = "Base Stat Total by Generation",
    x = "Generation", y = "Total BST"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

### 6.3 Stats by Primary Type (Top Types)

```{r stats-by-type}
# Get top 10 types
top_types <- pokemon_df |>
  count(type_1, sort = TRUE) |>
  head(10) |>
  pull(type_1)

pokemon_df |>
  filter(type_1 %in% top_types) |>
  ggplot(aes(x = reorder(type_1, total, median), y = total, fill = type_1)) +
  geom_boxplot(alpha = 0.7) +
  coord_flip() +
  labs(
    title = "BST by Primary Type (Top 10 Most Common)",
    x = "Primary Type", y = "Total BST"
  ) +
  theme(legend.position = "none")
```

## 7. Type Analysis

### 7.1 Type Distribution

```{r type-counts}
# Primary type
p1 <- pokemon_df |>
  count(type_1, sort = TRUE) |>
  head(10) |>
  ggplot(aes(x = reorder(type_1, n), y = n, fill = type_1)) +
  geom_col(alpha = 0.8) +
  coord_flip() +
  labs(title = "Top 10 Primary Types", x = "Type", y = "Count") +
  theme(legend.position = "none")

# Dual-type proportion
p2 <- pokemon_df |>
  count(is_dual_type) |>
  mutate(label = if_else(is_dual_type, "Dual-type", "Mono-type")) |>
  ggplot(aes(x = "", y = n, fill = label)) +
  geom_col(width = 1) +
  coord_polar("y") +
  labs(title = "Mono-type vs Dual-type", fill = "") +
  theme_void()

if (has_patchwork) {
  p1 | p2
} else {
  print(p1)
  print(p2)
}
```

### 7.2 Type Combinations

```{r type-combos}
# Top type combinations
pokemon_df |>
  filter(!is.na(type_2)) |>
  count(type_1, type_2, sort = TRUE) |>
  head(15) |>
  mutate(combo = paste(type_1, "/", type_2)) |>
  ggplot(aes(x = reorder(combo, n), y = n, fill = type_1)) +
  geom_col(alpha = 0.8) +
  coord_flip() +
  labs(
    title = "Top 15 Type Combinations",
    x = "Type Combination", y = "Count", fill = "Primary Type"
  )
```

### 7.3 Average Stats by Type

```{r avg-stats-type}
# Calculate average stats per type
type_stats <- pokemon_df |>
  group_by(type_1) |>
  summarise(
    count = n(),
    avg_total = mean(total, na.rm = TRUE),
    avg_attack = mean(attack, na.rm = TRUE),
    avg_defense = mean(defense, na.rm = TRUE),
    avg_speed = mean(speed, na.rm = TRUE)
  ) |>
  arrange(desc(avg_total))

print(kable(head(type_stats, 10), caption = "Top 10 Types by Average BST", digits = 1))

# Heatmap
type_stats |>
  head(12) |>
  select(type_1, avg_attack, avg_defense, avg_speed) |>
  pivot_longer(-type_1, names_to = "stat", values_to = "value") |>
  mutate(stat = str_remove(stat, "avg_")) |>
  ggplot(aes(x = stat, y = reorder(type_1, value), fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "darkred") +
  labs(
    title = "Average Stats Heatmap (Top 12 Types by BST)",
    x = "Stat", y = "Primary Type", fill = "Average Value"
  )
```

## 8. Special Categories Deep Dive

### 8.1 Legendary vs Regular Stats

```{r legendary-comparison}
# Statistical test
legendary_test <- t.test(total ~ is_legendary, data = pokemon_df)

cat("T-test: Legendary vs Regular BST\n")
cat("p-value:", legendary_test$p.value, "\n")
cat("Mean Regular:", legendary_test$estimate[1], "\n")
cat("Mean Legendary:", legendary_test$estimate[2], "\n\n")

# Violin plot
pokemon_df |>
  mutate(label = if_else(is_legendary, "Legendary", "Regular")) |>
  ggplot(aes(x = label, y = total, fill = label)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.2, alpha = 0.5) +
  scale_fill_manual(values = c("Regular" = "steelblue", "Legendary" = "gold")) +
  labs(
    title = "BST: Legendary vs Regular Pokemon",
    x = "", y = "Total BST"
  ) +
  theme(legend.position = "none")
```

### 8.2 All Special Categories

```{r special-categories}
category_stats <- pokemon_df |>
  group_by(category) |>
  summarise(
    count = n(),
    avg_total = mean(total, na.rm = TRUE),
    avg_hp = mean(hp, na.rm = TRUE),
    avg_attack = mean(attack, na.rm = TRUE),
    avg_defense = mean(defense, na.rm = TRUE),
    avg_sp_atk = mean(sp_atk, na.rm = TRUE),
    avg_sp_def = mean(sp_def, na.rm = TRUE),
    avg_speed = mean(speed, na.rm = TRUE)
  ) |>
  arrange(desc(avg_total))

print(kable(category_stats, caption = "Average Stats by Category", digits = 1))
```

## 9. Generation Evolution Analysis

### 9.1 Power Creep Over Generations

```{r generation-power-creep}
gen_stats <- pokemon_df |>
  group_by(generation) |>
  summarise(
    count = n(),
    avg_total = mean(total, na.rm = TRUE),
    median_total = median(total, na.rm = TRUE),
    sd_total = sd(total, na.rm = TRUE)
  )

ggplot(gen_stats, aes(x = generation)) +
  geom_line(aes(y = avg_total, group = 1, color = "Average"), size = 1.2) +
  geom_line(aes(y = median_total, group = 1, color = "Median"), size = 1.2) +
  geom_point(aes(y = avg_total), size = 3) +
  geom_point(aes(y = median_total), size = 3) +
  scale_color_manual(values = c("Average" = "blue", "Median" = "red")) +
  labs(
    title = "Average & Median BST by Generation",
    x = "Generation", y = "BST", color = "Metric"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 9.2 Type Diversity by Generation

```{r type-diversity}
pokemon_df |>
  group_by(generation, type_1) |>
  summarise(count = n(), .groups = "drop") |>
  ggplot(aes(x = generation, y = count, fill = type_1)) +
  geom_col(position = "stack") +
  labs(
    title = "Type Distribution by Generation",
    x = "Generation", y = "Count", fill = "Primary Type"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 10. Key Findings Summary

```{r findings}
library(glue)

cat(glue("=== EDA KEY FINDINGS ===

1. Dataset Overview:
   - Total Pokemon: {nrow(pokemon_df)}
   - Legendary: {sum(pokemon_df$is_legendary)}
   - Mythical: {sum(pokemon_df$is_mythical)}
   - Regular: {sum(!pokemon_df$is_special)}

2. Base Stats:
   - Mean BST: {round(mean(pokemon_df$total, na.rm = TRUE), 1)}
   - BST Range: {min(pokemon_df$total, na.rm = TRUE)} - {max(pokemon_df$total, na.rm = TRUE)}

3. Type Distribution:
   - Unique primary types: {length(unique(pokemon_df$type_1))}
   - Dual-type Pokemon: {sum(pokemon_df$is_dual_type)} ({round(100 * mean(pokemon_df$is_dual_type), 1)}%)

4. Statistical Tests:
   - Legendary vs Regular BST p-value: {format(legendary_test$p.value, scientific = TRUE, digits = 3)}
   - Difference is {if_else(legendary_test$p.value < 0.001, 'highly', '')} significant

5. Correlations:
   - Total vs Attack: {round(cor(pokemon_df$total, pokemon_df$attack, use = 'complete.obs'), 3)}
   - Attack vs Defense: {round(cor(pokemon_df$attack, pokemon_df$defense, use = 'complete.obs'), 3)}
   - Height vs Weight: {round(cor(pokemon_df$height_m, pokemon_df$weight_kgs, use = 'complete.obs'), 3)}

6. Physical Attributes:
   - BMI Range (excluding outliers): {round(min(pokemon_df$bmi[pokemon_df$bmi < 1000], na.rm = TRUE), 1)} - {round(max(pokemon_df$bmi[pokemon_df$bmi < 1000], na.rm = TRUE), 1)}
   - Extreme outlier: Cosmoem (BMI = 99,990) due to ultra-dense core
   - Height range: {min(pokemon_df$height_m, na.rm = TRUE)} - {max(pokemon_df$height_m, na.rm = TRUE)} m
   - Weight range: {min(pokemon_df$weight_kgs, na.rm = TRUE)} - {max(pokemon_df$weight_kgs, na.rm = TRUE)} kg
"))
```

## 11. Export Enhanced Dataset

```{r export}
# Save enriched dataset with all features
if (!dir.exists("data")) dir.create("data", recursive = TRUE)

write_csv(pokemon_df, "data/pokemon_data_enriched.csv")

cat("\nEnriched dataset saved to: data/pokemon_data_enriched.csv\n")
cat(
  "New features added:",
  "is_legendary, is_mythical, is_paradox, is_ultra_beast,",
  "category, generation, type one-hot columns\n"
)
```

---

