---
title: "Data Source and Data Cleaning Process"
output: html_document
---
```{r}
library(rvest)
library(tidyverse)
```


### Data Source  
The data used in this project were scraped from pokemondb.net, a resource that maintains one of the most complete collections of Pokémon information to date, encompassing all nine generations (1,025 Pokémon species including alternate forms).  
The dataset includes variables such as base statistics, physical dimensions (height and weight), and type attributes.  
A portion of the metadata is sourced from the Veekun database, with additional details compiled by community contributors based on in-game data.


```{r}
raw_state_df <- read_html("data/Pokemon.html") |> 
  html_table()


raw_wh_df <- read_html("data/Pokemon by height and weight.html") |> 
  html_table()
```

```{r}
pokemon_state_df <- as_tibble(raw_state_df[[1]]) |> 
  janitor::clean_names()

pokemon_df <- as_tibble(raw_wh_df[[1]]) |> 
  janitor::clean_names() |> 
  right_join(pokemon_state_df, by = c("name", "type", "number"))
```

